---
title: "Binary Logistic Regression: Florida Palmetto "
description: |
  Use `ggpairs` to explore multivariate relationships for Florida palmetto species and use binary logistic regression to test the feasability of using different variables to determine if a palmetto is a certain species. 
author:
  - name: Katelin Seeto
    url: {}
date: 03-12-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}

library(tidyverse)
library(here)
library(GGally) 
library(broom) 
library(jtools)
library(knitr)
library(kableExtra)
library(patchwork)

```

```{r}
# Read in and wrangle data

palmetto <- read_csv(here("_posts", "2021-03-12-binarylogisticregression", "data", "palmetto.csv"),
                     col_types = cols(.default = 'c')) %>% 
  mutate(height = as.numeric(height),
         length = as.numeric(length),
         width = as.numeric(width),
         green_lvs = as.numeric(green_lvs),
         species = recode(species, "1" = "Serenoa repens","2" = "Sabal etonia"))
```

### **Data visualizations**
```{r}
# Get an initial broad idea of correlations, distributions, and relationships

# palmetto %>% 
#   select(species, height, length, width, green_lvs) %>% 
#   ggpairs(aes(color = species))

# set colors for species
colors <- c("olivedrab3", "green4")
```
 
```{r, fig.cap= "**Fig.1:** A comparison of the plant height (cm), and the widest length of the canopy (cm) for *Sabal etonia* (light green), and *Serenoa repens* (dark green), shows *Sabal etonia* plants have a slightly larger canopy length, while *Serenoa repens* plants tend to be slightly taller."}

#height
h <- palmetto %>% 
  ggplot(aes(x = species, y = height,
             color = species)) + 
  geom_jitter()+
  scale_color_manual(values = colors)+
  labs(y = "Plant height (cm)")+
  theme_minimal()+
  theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.text.x = element_blank())

#length
l <- palmetto %>% 
  ggplot(aes(x = species, y = length,
             color = species)) + 
  geom_jitter()+
  scale_color_manual(values = colors)+
  theme_minimal()+
  labs(y = "Canopy length (cm)")+
  theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.text.x = element_blank())

#length vs. height 
l_vs_h <- palmetto %>% 
  ggplot(aes(x = length, y = height,
             color = species)) + 
  geom_point()+
  scale_color_manual(values = colors)+
  labs(x = "Canopy length (cm)",
       y = "Plant height (cm)",
       color = "Species")+
  facet_wrap(species~.)+
  theme_minimal()+
  theme(strip.text = element_blank())

l_vs_h +(l/h) +
  plot_layout(guides = 'collect',
              ncol = 2, width = c(2,1))

```


```{r, fig.cap= "**Fig.2:** A comparison of the number of green leaves on a plant, and the widest width of the canopy (cm) for *Sabal etonia* (light green), and *Serenoa repens* (dark green), shows that both species are similar in width, but *Serenoa repens* appears to have a greater number of green leaves."}

# width
w <- palmetto %>% 
  ggplot(aes(x = species, y = width,
             color = species)) + 
  geom_jitter()+
  scale_color_manual(values = colors)+
  theme_minimal()+
  labs(y = "Canopy width (cm)")+
  theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.text.x = element_blank())

# green leaves
g <- palmetto %>% 
  ggplot(aes(x = species, y = green_lvs,
             color = species)) + 
  geom_jitter()+
  scale_color_manual(values = colors)+
  theme_minimal()+
  labs(y = "Number of green leaves")+
  theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.text.x = element_blank())

# width vs. green leaves 
w_vs_g <- palmetto %>% 
  ggplot(aes(x = width, y = green_lvs,
             color = species)) + 
  geom_point()+
  scale_color_manual(values = colors)+
  labs(x = "Canopy width (cm)",
       y = "Number of green leaves",
       color = "Species")+
  facet_wrap(species~.)+
  theme_minimal()+
  theme(strip.text = element_blank())

w_vs_g +(w/g) +
  plot_layout(guides = 'collect',
              ncol = 2, width = c(2,1))

````



### **Binary logistic regression**

**Table 1** Binary logistic regression model results using maximum height, widest length of canopy, widest width of canopy, and  count of green leaves, as predictor variables to determin the probability that a plant is *Serenoa repens*. 
```{r results='asis'}

# Change species to factor level and then check levels 
palmetto <- palmetto %>% 
  mutate(species = as.factor(species))
# levels (palmetto$species)
# 0 = Sabal etonia, 1 = Serenoa repens


# Binary logistic regression model to see relationship between: maximum height, widest length of canopy, widest width of canopy, and  count of green leaves, as potential  predictors of palmetto species  
palmetto_blr <- glm(species ~ height + width + length + green_lvs,
                    data = palmetto, 
                    family = "binomial")

#summary(palmetto_blr)

# Get tidy version with broom
blr_tidy <- broom::tidy(palmetto_blr) %>% 
  mutate(p.value = case_when(p.value < 0.0001 ~ "<0.0001",# p values <0.0001 will all be labeled "<0.0001"
                             TRUE ~ as.character(p.value)),
         )
  
kable(blr_tidy,
      col.names = c("Term",
                    "Coefficient",
                    "Standard Error", 
                    "Statistic",
                    "p-value"),
      align = 'c') %>% 
  kable_styling()
  

```
### **Model Evaluation** 

**Table 2** Evaluation of the binary logistic regression model's success rate for species identification of *Serenoa repens* and *Sabal etonia* by number of plants correctly, and incorrectly classified, and percentage of plants that were classified correctly by the model.
```{r}
blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response") %>% 
  mutate(predicted_species = case_when(.fitted >= 0.5 ~ "Serenoa repens",
                                       .fitted <= 0.5 ~ "Sabal etonia"),
         correct= if_else(species == predicted_species, "Correct", "Incorrect"))

# Table showing how successful model was at predicting species
blr_eval <- blr_fitted %>% 
  group_by(species) %>% 
  count(correct) %>% 
  pivot_wider(names_from = correct, values_from = n) %>% 
  mutate(percent_correct = round((Correct/sum(Correct, Incorrect)*100),2)) 
kable(blr_eval, 
      col.names = c("Species",
                    "Classified Correctly",
                    "Classified Incorrectly",
                    "% Classified Correctly "),
       align = 'c') %>% 
  kable_styling()
  

```


#### Data Citation

Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5





