---
title: "Finding an equation with Parameters"
description: |
  A short description of the post.
author:
  - name: Katelin Seeto
    url: {}
date: 03-12-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}

# Attach packages 
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(nlstools)
library(gt)
library(latex2exp)

```

```{r}
# Read in global wild fish catch data 
fish <- read_csv(here("_posts", "2021-03-12-parameters", "data", "fish_catch.csv"),
                 skip = 2,
                 col_names = TRUE)

# Tidy and clean data 
fish_clean <- fish %>% 
  clean_names() %>% 
  drop_na() %>% 
  mutate(across(c(year, wild_catch), as.numeric))

# Create df for wild catch with column for time in years, with 1950 = 0
wild <- fish_clean %>% 
  mutate(time = seq(from = 0, 
                    to = (nrow(fish_clean)-1),
                    by = 1)) %>% 
  select(time, wild_catch)

# Create exploratory graph of wild fish catch over time 
ggplot(fish_clean, aes(x = year, y = wild_catch)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1950,2012,10)) +
  scale_y_continuous(breaks = seq(15,100,10)) +
  labs(title = "Exploratory plot of wild fish catch over time",
       x = "year",
       y = "wild catch (million tons)") +
  theme_light()

# Estimate growth constant during the exponential phase to get an estimate for growth rate constant (k)

# Select only observations occurring during exponential growth phase (time <45 = year <1995), and log transform wild_catch
wild_exp <- wild %>% 
  filter(time < 45) %>% 
  mutate(ln_catch = log(wild_catch))

# Model linear to get estimate of *k*
lm_k <- lm(ln_catch ~ time, data = wild_exp)
# lm_k # ~ 0.03
  
```
The data trend can be described by logistic growth equation: 
$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant

My initial estimates for the parameters are:    

- $K$ :  ~**90**  
- $A$ : $\frac{90-17}{17}$ ~**4**  
- $k$ :  ~**0.03**  


```{r echo=FALSE}
# Find parameters for wild catch model using nonlinear least squares

# Enter model information with list of estimated starting parameter values into `nls()`
wild_nls <- nls(wild_catch ~K/(1 + A*exp(-k*time)),
                data = wild, 
                start = list(K = 90,
                             A = 4, 
                             k = 0.03))

# View model summary 
# summary(wild_nls)

# Get model in tidier format using broom::tidy
model_out <- broom::tidy(wild_nls)

# Create table of parameter outcomes 
model_out %>% 
  select(term, estimate) %>% 
  gt() %>% 
  tab_header(title = md("Nonlinear Least-squares Parameter<br/> Estimates for Wild Fish Catch Model")) %>% 
  tab_options(table.width = pct(50)) %>% 
  tab_style(style = list(
    cell_text(align = "center",
              weight = "bold")),
    locations = cells_column_labels(columns = vars(term, estimate))) %>%
  tab_style(style = list(
    cell_text(align = "center")),
    locations = cells_body(columns = vars(estimate))) %>%
  tab_style(style = list(
    cell_text(style = "italic",
              align = "center")),
    locations = cells_body(columns = vars(term))) %>%
  fmt_number(columns = vars(estimate),
             decimals = 2) %>% 
  cols_label(term = "Parameter",
             estimate = md("Estimate<br/> (catch in million tons)"))
```

```{r}
# Make predictions for global wild catch for each year (time) in the original data frame
p_predict <- predict(wild_nls)

# Bind predictions to original data frame
wild_complete <- data.frame(wild, p_predict)

# Add column listing years back in for easier plot labeling 
wild_complete$year <- fish_clean$year

# Plot predictions from model output against original data 
ggplot(data = wild_complete, aes(x = year, y = wild_catch)) + 
  geom_point(color = "cadetblue", 
             size = 2) + 
  geom_line(aes(x = year, y = p_predict), 
            color = "orangered",
            size = 1) +
  annotate("text",
           label = "Model Prediction",
             x = 2020, 
             y = 97,
             color = "orangered",
             size = 3.5) +
  annotate("text",
           label = "Catch Data",
             x = 2020, 
             y = 91,
             color = "cadetblue",
             size = 3.5) +
  annotate("text", 
           label = TeX("$P(t) = \\frac{100.28}{1 + 4.32e^{-0.07t}}$"), 
           size = 4, 
           color = "orangered",
           x = 2008, 
           y = 10) +
  coord_cartesian(clip = "off") +
  scale_x_continuous(breaks = seq(1950,2012,10),
                     expand = c(0,1)) +
  scale_y_continuous(limits = c(0,100), 
                     breaks = seq(10,100, 10),
                     expand = c(0,0)) +
  labs(x = "Year",
       y = "Global wild fish catch\n(million tons)") + 
  theme_bw() +
  theme(axis.line = element_line(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(t= 0.5, r = 4, b = 0.5, l = 0.5, unit = "cm"))
```


### Data source:
 Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.


